---
title: "Question 2, Unknown Standard Deviation"
author: "Joseph Froelicher"
date: "10/1/2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, include = TRUE)
```

```{r}
A <- power.t.test(n=5, delta=100, sd=75, sig.level=0.05, power=NULL, type='one.sample')
A$power
```
Power for this unknown standard deviation case is 61.4%.
  
```{r}
B <- power.t.test(n=NULL, delta=100, sd=75, sig.level=0.05, power=0.9, type='one.sample')
B$n
```
The number of participant required for this study with unknown standard deviation is 8.07, which we always round up to 9 participants.
  
```{r}
C <- power.t.test(n=5, delta=NULL, sd=75, sig.level=0.05, power=0.9, type='one.sample')
C$delta
```
The smallest mean change in beta carotene required for this scenario with unknown standard deviation and 90% power is 147.4.  
```{r}
D <- power.t.test(n=5, delta=NULL, sd=75, sig.level=0.05, power=0.8, type='one.sample')
D$delta
```
The smallest mean change in beta carotene required for this scenario with unknown standard deviation 80% power is 126.1.
