---
title: "Question 2"
author: "Joseph Froelicher"
date: "10/15/2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, include = TRUE)

setwd("C:/Users/froelijo/dev/methods1/hw6")
procedure <- read.csv("ProcedureCost.csv")
head(procedure)
old <- procedure[procedure$Procedure == 1,]
new <- procedure[procedure$Procedure == 2,]
head(new)
head(old)
```
# Part A
```{r}
n1 <- dim(new)[1]
n2 <- dim(old)[1]

B <- 10000

boot_ratio <- vector("double", B)

set.seed(8675309)

for (i in 1:B) {
  val1 <- sample(new$Cost, n1, replace = TRUE)
  val2 <- sample(old$Cost, n2, replace = TRUE)
  
  boot_ratio[i] <- mean(val1) - mean(val2)
}

par(mfrow = c(1, 2))
hist(boot_ratio, main = "Bootstrap of Ratio", breaks = 50, xlab = 'difference of means(new - old)')
qqnorm(boot_ratio)
qqline(boot_ratio)
```
  
# Part B
The histogram of the bootstrap data seem to be bell-shaped, with very few outliers from the normal quantile line on the normal QQ plot. 

# Part C
```{r}
boot_mean <- mean(boot_ratio)
boot_bias <- boot_mean - (mean(new$Cost) - mean(old$Cost))
boot_se <- sd(boot_ratio)
```

# Part D
```{r}
alpha = 0.05

normal_lower <- boot_mean - qnorm(1 - (alpha / 2)) * boot_se
normal_upper <- boot_mean + qnorm(1 - (alpha / 2)) * boot_se

coverage_lower <- sum(boot_ratio < normal_lower) / B
coverage_upper <- sum(boot_ratio > normal_upper) / B

boot_lower <- quantile(boot_ratio, 0.025)
boot_upper <- quantile(boot_ratio, 0.975)

accuracy <- boot_bias / boot_se
```

Based on our estimates of coverage [`r coverage_lower`, `r coverage_upper`], the 95% normal percentile estimates [`r normal_lower`, `r normal_upper`] are pretty good for both the lower and upper bounds. The lower bound has coverage of `r coverage_lower * 100`%, and the upper bound has coverage of `r coverage_upper * 100`%, these are close to our target of 2.5%. The 95% bootstrap confidence interval is [`r boot_lower`, `r boot_upper`]. The accuracy of the bootstrap confidence interval is `r accuracy`, which is less than the suggested cutoff of 0.10, indicating good accuracy.