---
title: "Question 2"
author: "Joseph Froelicher"
date: "December 3, 2020"
output: pdf_document

---
```{r, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE, include = TRUE)
setwd("C:/Users/froelijo/dev/methods1/hw10")
lead = read.csv('lead2.csv')
```

# Part A
```{r, echo = TRUE, include = TRUE}
fit1 = lm(iq ~ expose, data = lead)
fit2 = lm(iq ~ expose + resdur, data = lead)
fit3 = lm(resdur ~ expose, data = lead)

summary(fit1)
summary(fit2)
summary(fit3)

gamma_x = fit3$coefficients[2] # gamma_x
beta_m = fit2$coefficients[3] # beta_mediator
beta_adj = fit2$coefficients[2] # beta adjusted
beta_crude = fit1$coefficients[2] # beta crude
```
$\gamma_x\:\:\:\:\:\:\:|\:\:\:\:\:\beta_{adj}\:\:\:\:\:|\:\:\:\:\beta_{crude}\:\:\:\:|\:\:\:\:\:\:\:\beta_m$\newline
`r gamma_x` | `r beta_m` | `r beta_adj` | `r beta_crude`
\newline
\flushleft
  
# Part B
IQ was mediated `r (fit1$coefficients[2] - fit2$coefficients[2]) / fit1$coefficients[2] * 100`% by duration at residence.
\newline

# Part C
```{r, echo = TRUE, include = TRUE}
se = sqrt(
  ( ( fit3$coefficients[2] ^ 2 ) *
  ( summary(fit2)$coefficients[3, 2] ^ 2 ) ) +
  ( ( fit2$coefficients[3] ^ 2 ) *
  ( summary(fit3)$coefficients[2, 2] ^ 2 ) )
)

z = (fit1$coefficients[2] - fit2$coefficients[2]) / se
p = 2 * pnorm(z)
alpha = 0.05

confint = c(
  ( ( fit1$coefficients[2] - fit2$coefficients[2] ) - ( qnorm(1 - (alpha / 2)) * se ) ) / fit1$coefficients[2] * 100,
  ( ( fit1$coefficients[2] - fit2$coefficients[2] ) + ( qnorm(1 - (alpha / 2)) * se ) ) / fit1$coefficients[2] * 100
)
```
 The 95% confidence interval for percent mediated by duration at residence using the normal approximation to estimate standard error is [`r confint[2]`, `r confint[1]`] with p = `r p`.
\newline
 
# Part D
```{r, echo = TRUE, include = TRUE}
n <- dim(lead)[1]
b <- 10000

se_vec = vector("double", b)
fit_mat = matrix(NA, nrow = b, ncol = 6)

set.seed(8675309)

for (i in 1:b) {
  new = lead[sample(nrow(lead), n, replace = TRUE), ]
  
  fit1 = lm(iq ~ expose, data = new)
  fit2 = lm(iq ~ expose + resdur, data = new)
  fit3 = lm(resdur ~ expose, data = new)
  
  fit_mat[i, 1] = fit3$coefficients[2] # gamma_x
  fit_mat[i, 2] = summary(fit2)$coefficients[3, 2] # se of beta_mediator
  fit_mat[i, 3] = fit2$coefficients[3] # beta_mediator
  fit_mat[i, 4] = summary(fit3)$coefficients[2, 2] # se of gamma_x
  fit_mat[i, 5] = fit1$coefficients[2] - fit2$coefficients[2] # indirect effect
  fit_mat[i, 6] = fit1$coefficients[2] # total effect
}

se_boot = sqrt(
  ( mean(fit_mat[, 1]) ^ 2 ) *
  ( mean(fit_mat[, 2]) ^ 2 ) +
  ( mean(fit_mat[, 3]) ^ 2 ) *
  ( mean(fit_mat[, 4]) ^ 2 )
)

boot_confint = c(
  ( mean(fit_mat[, 5]) - ( qnorm(1 - (alpha / 2)) * se_boot ) ) / mean(fit_mat[, 6]) * 100,
  ( mean(fit_mat[, 5]) + ( qnorm(1 - (alpha / 2)) * se_boot ) ) / mean(fit_mat[, 6]) * 100
)

z_boot = mean(fit_mat[, 5]) / se_boot
p_boot = 2 * pnorm(z_boot)
```
The 95% bootstrap confidence interval for the percent mediated is [`r boot_confint[2]`, `r boot_confint[1]`], with p = `r p_boot`. This is very similar to what we saw from part C, but as we would expect, it is slightly narrower.