---
title: "Question 3"
author: "Joseph Froelicher"
date: "11/19/2020"
output: pdf_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE, include = TRUE)

setwd("C:/Users/froelijo/dev/methods1/hw9")

carotenoids <- read.table('carotenoids.dat')

colnames(carotenoids) <- c(
  'age',
  'sex',
  'smoke',
  'bmi',
  'vitamins',
  'calories',
  'fat',
  'fiber',
  'alcohol',
  'chol',
  'betadiet',
  'retdiet',
  'betaplas',
  'retplas'
)
```

# Part A
```{r}
never_mean <- mean(carotenoids[carotenoids$smoke == 1,]$betaplas)
never_sd <- sd(carotenoids[carotenoids$smoke == 1,]$betaplas)
never_se <- never_sd / sqrt(length(carotenoids[carotenoids$smoke == 1,]$betaplas))

former_mean <- mean(carotenoids[carotenoids$smoke == 2,]$betaplas)
former_sd <- sd(carotenoids[carotenoids$smoke == 2,]$betaplas)
former_se <- former_sd / sqrt(length(carotenoids[carotenoids$smoke == 2,]$betaplas))

current_mean <- mean(carotenoids[carotenoids$smoke == 3,]$betaplas)
current_sd <- sd(carotenoids[carotenoids$smoke == 3,]$betaplas)
current_se <- current_sd / sqrt(length(carotenoids[carotenoids$smoke == 3,]$betaplas))

data.frame(
  "Never" = c(never_mean, never_sd, never_se),
  "Former" = c(former_mean, former_sd, former_se),
  "Current" = c(current_mean, current_sd, current_se),
  row.names = c("mean", "sd", "se")
)
```
  
# Part B
```{r}
carotenoids$never <- carotenoids$smoke == 1
carotenoids$former <- carotenoids$smoke == 2
carotenoids$current <- carotenoids$smoke == 3

fit_ref <- glm(betaplas ~ former + current + calories, data = carotenoids)
```
$$Y_{betaplas} = \beta_{never} + \beta_{former}X_{1} + \beta_{current}X_2 + \beta_{calories}X_3 + \epsilon_i$$
Where the intercept $\beta_0$ is the expected mean for never smokers. We are assuming the residuals to be normally distributed.
  
# Part C
```{r}
fit_empty <- glm(betaplas ~ 1, data = carotenoids)
aov_empty <- anova(fit_ref, fit_empty, test = "F")
```
$$H_0: \beta_{never} = \beta_{former} = \beta_{current} = \beta_{calories} = 0$$
$$H_A:  \beta_{never}\neq0\:or\:\beta_{former}\neq0\:or\:\beta_{current}\neq0\:or\:\beta_{calories}\neq0$$
Fail to reject the null hypothesis that smoking status and calories do not contribute significantly to the prediction of Beta-Carotine levels (p = `r aov_empty[2,6]`).

# Part D
```{r}
fit_smoke <- glm(betaplas ~ former + current, data = carotenoids)
aov_smoke <- anova(fit_smoke, fit_empty, test = "F")
```
$$H_0: \beta_{never} = \beta_{former} = \beta_{current} = 0$$
$$H_A:  \beta_{never}\neq0\:or\:\beta_{former}\neq0\:or\:\beta_{current}\neq0$$
Reject the null hypothesis that smoking status does not contribute significantly to the prediction of Beta-Carotine levels (p = `r aov_smoke[2,6]`).
  
# Part E
```{r}
aov_ref <- anova(fit_ref, fit_smoke, test = "F")
```
Fail to reject the null hypothesis that calories contributes significantly to the prediction of Beta-Carotine levels (p = `r aov_ref[2,6]`)
