---
title: "Question 2"
author: "Joseph Froelicher"
date: "11/19/2020"
output: pdf_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE, include = TRUE)
options(scipen = 999)

library(DescTools)

alpha <- 0.05

lung <- data.frame(
  group=c(
    rep('A',5),
    rep('B',12),
    rep('C',5)
  ),
  react=c(
    20.8,4.1,30,24.7,13.8,7.5,7.5,11.9,4.5,3.1,8,
    4.7,28.1,10.3,10,5.1,2.2,9.2,2,2.5,6.1,7.5
  )
)
```

# Part A
```{r}
oneway_test <- oneway.test(react ~ group, data = lung, var.equal = TRUE)
```
$$H_0: \mu_a = \mu_b = \mu_c$$
$$H_0: \mu_a\neq\mu_b\:or\:\mu_a\neq\mu_c,\:or\:\mu_b\neq\mu_c$$

We reject the null hypothesis that there is no mean difference between groups A, B and C ($\alpha$ = `r alpha`. At least one group mean is different from another.

# Part B
```{r}
aov <- aov(react ~ group, data = lung)
hsd <- PostHocTest(aov, method = "hsd")
p1 <- hsd$group[1,4]
p2 <- hsd$group[2,4]
```
  
After comparing the groups, two p-values were significant, (A to B and A to C, with $\alpha$ = `r alpha`), and with the number of groups being at least 3 (number of comparisons at least 3) then we should perform a post-hoc Tukey-HSD test for multiple comparisons. After performing a post-hoc Tukey's HSD test, group A is still significantly different than group B (p = `r p1`), and group A is still significantly different than group C p = `r p2`)

# Part C
```{r}
oneway_test <- oneway.test(react ~ group, data = lung, var.equal = FALSE)
```
  
In this case it is not necessary to do any more post-hoc analyses, the p-value is non-significant, post-hoc analyses will not make our result significant.