---
title: "Question 2"
author: "Joseph Froelicher"
date: "10/8/2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, include = TRUE)

```
```{r}
nausea <- data.frame(
  c('A', 'B'),
  "Yes" = c(65, 72),
  "No" = c(15, 48),
  row.names = 1
)

nausea

nausea_table <- as.table(cbind(c(65, 72), c(15, 48)))
nausea_table
```
# Part A
```{r}
# by hand
alpha <- 0.05

# risk difference
rd <- (nausea[1, 1] / ( nausea[1, 1] + nausea[1, 2] )) - (nausea[2, 1] / ( nausea[2, 1] + nausea[2, 2]) )

rd_ci <- c(
  rd + (
    qnorm(1 - (alpha / 2)) * sqrt(
    (( nausea[1, 1] + nausea[2, 1] ) / ( nausea[1, 1] + nausea[2, 1] + nausea[1, 2] + nausea[2, 2] )) *
      ( 1 - ((nausea[1, 1] + nausea[2, 1]) / (nausea[1, 1] + nausea[2, 1] + nausea[1, 2] + nausea[2, 2])) ) *
      ( (nausea[1, 1] + nausea[1, 2])^-1 + (nausea[2, 1] + nausea[2, 2])^-1 )
    )
  ),
  exp(log(rd)) - (
    qnorm(1 - (alpha / 2)) * sqrt(
    (( nausea[1, 1] + nausea[2, 1] ) / ( nausea[1, 1] + nausea[2, 1] + nausea[1, 2] + nausea[2, 2] )) *
      ( 1 - ((nausea[1, 1] + nausea[2, 1]) / (nausea[1, 1] + nausea[2, 1] + nausea[1, 2] + nausea[2, 2])) ) *
      ( (nausea[1, 1] + nausea[1, 2])^-1 + (nausea[2, 1] + nausea[2, 2])^-1 )
    )
  )
)

# risk ratio
rr <- ( nausea[1, 1] / ( nausea[1, 1] + nausea[1, 2] ) ) / ( nausea[2, 1] / ( nausea[2, 1] + nausea[2, 2] ) )

rr_ci <- c(
  exp(
    log(rr) + (
      qnorm(1  - (alpha / 2)) * sqrt(
        nausea[1, 2] / ( nausea[1, 1] * ( nausea[1, 1] + nausea[1, 2] ) ) +
        nausea[2, 2] / ( nausea[2, 1] * ( nausea[2, 1] + nausea[2, 2] ) )
      )
    )
  ),
  exp(
    log(rr) - (
      qnorm(1 - (alpha / 2)) * sqrt(
        nausea[1, 2] / ( nausea[1, 1] * ( nausea[1, 1] + nausea[1, 2] ) ) +
        nausea[2, 2] / ( nausea[2, 1] * ( nausea[2, 1] + nausea[2, 2] ) )
      )
    )
  )
)

# odds ratio
or <- ( nausea[1, 1] * nausea[2, 2] ) / ( nausea[2, 1] * nausea[1, 2] )

or_ci <- c(
  exp(
    log(or) + (
      qnorm(1 - (alpha / 2)) * sqrt(nausea[1, 1]^-1 + nausea[1, 2]^-1 + nausea[2, 1]^-1 + nausea[2, 2]^-1)
    )
  ),
  exp(log(or) - (
      qnorm(1 - (alpha / 2)) * sqrt(nausea[1, 1]^-1 + nausea[1, 2]^-1 + nausea[2, 1]^-1 + nausea[2, 2]^-1)
    )
  )
)

data.frame(
  'statistic' = c('rd', 'rr', 'or'),
  'point_estimate' = c(rd * 100, rr, or),
  'lower' = c(rd_ci[2] * 100, rr_ci[2], or_ci[2]),
  'upper' = c(rd_ci[1] * 100, rr_ci[1], or_ci[1])
)

# using epiR
library(epiR)

epi.2by2(nausea_table)
```
There were `r rd`(`r rd_ci[2]`, `r rd_ci[1]`) more subjects in the anesthetic group A compared to the anesthetic group B. Notice the discrepency between the hand calculation CI and the epiR function generated CI. This is due to a slightly different adjustment for the statistics at hand. Those who received anesthetic A, were `r rr`(`r rr_ci[2]`, `r rr_ci[1]`) times more likely to experience nausea. And, the odds of nausea in those using anesthetic A is `r or`(`r or_ci[2]`, `r or_ci[1]`) times the odds in those using anesthetic B. All three of these estimates agree with the significants implied by their confidence intervals.

# Part B
```{r}
chisq <- chisq.test(nausea_table, correct = FALSE)
chisq_correct <- chisq.test(nausea_table, correct = TRUE)
fisher <- fisher.test(nausea_table)
mcnemar <- mcnemar.test(nausea_table)

df <- data.frame(
  'approach' = c('chisq.test', 'chisq.test.corrected', 'fisher.test', 'mcnemar.nest'),
  'test_statistic' = c(chisq$statistic, chisq_correct$statistic, NA, mcnemar$statistic),
  'p_value' = round(c(chisq$p.value, chisq_correct$p.value, fisher$p.value, mcnemar$p.value), digits = 3)
)

iter <- 1
for (i in df$p_value) {
  if (i < 0.001) {
    df$p_value[iter] <- '<0.001'
  }
  
  iter <- iter + 1
}

df
```

# Part C
McNemar's test was originally intended to be used for paired data. There is no indication that any pairing was done between the two groups anesthetic A and anesthetic B. Additionally, I tend to lean towards more conservative estimates, so the corrected chi-square test will be a more conservative estimate that the chi-square test with no correction, and likely a better test. At the very least we don't need to do both.

# Part D
The most appropriate test here is likely the corrected chi-square test. While the Fisher's exact test is a good test for low sample size, it seems as though the correction on chi-squared tests is creating a more conservative test. That being said, the chi-square test with a correction tells us that with 95% confidence we can say that the use of anesthetic A implies a significant result of nausea.